---
title: "Data Analysis and Visualization"
author: "Victoria Oldemburgo de Mello"
date: "10/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## The research project
In this research project, we wanted to understand how **social media use** is related to **well-being** and **political polarization**.

Because researchers have failed to find relationships between overall social media use (measured in time) and the dependent variables, I decided to approach the question such that social media use could be divided into different types of uses. I wanted to investigate, for instance, if using social media to connect with friends would be any different from using social media to check for news.

To measure the different types of social media use, I created a scale "functions of social media use". After performing a factor analysis, the scale ended up divided into five factors:
- Interacting: Using social media to connect with a community, sharing information with people. 
- Information Seeking: Using social media for shopping, finding information about businesses, news consumption.
- Relaxation: Using social media to distract oneself from stressful events, as a coping mechanism.
- Entertainment: Using social media to find fun content, to have fun.
- Commercial: Using social media to promote one's own brand or oneself. 

## Hypotheses
We had 4 main hypotheses to test in this study.

#### 1) Individual traits will be linked to social media use.
We expect that personality traits will be related to different functions of social media use.

#### 2) Using social media for news consumption will be positively associated with polarization.

#### 3) Dispositional states will be related to social media.
We expect that boredom, loneliness, and self-esteem will be related to social media use. 

#### 4) Using social media actively or passively will be associated with well-being.
We expect that using social media actively will hold a positive relationship with well-being while using it passively will hold a negative relationship.



```{r, echo=FALSE}
library(tidyverse)

setwd('C:/Users/Artur/OneDrive - University of Toronto/Desktop/thesis -docs/Study 1')
df <- read.csv("clean_dataset.csv")

df1 <- df # keeping the original df intact in case I need it later
```

### Investigating First hypothesis: Individual Traits x Social Media Use
Because we have 297 observations, all correlations larger than 0.114 will be significant (p<=0.05).
To simplify, we will only run the estimate parameter of each correlation.

```{r}
##### EXTRAVERSION AND FUNCTIONS: CORRELATIONS #####
cor.test(df1$extraversion, df1$interacting)$estimate # sig
cor.test(df1$extraversion, df1$commercial)$estimate  # sig
cor.test(df1$extraversion, df1$relaxation)$estimate  #N.S
cor.test(df1$extraversion, df1$entertainment)$estimate  #N.S
cor.test(df1$extraversion, df1$info_seeking)$estimate  #sig

##### EXTRAVERSION AND FUNCTIONS: VISUALIZING THE SIGNIFICANT RELATIONSHIPS #####
fig1 <- ggplot(df1, aes(x=extraversion, y=interacting))
fig1 + geom_point() + geom_smooth(method = "lm")

fig2 <- ggplot(df1, aes(x=extraversion, y=commercial))
fig2 + geom_point() + geom_smooth(method = "lm")

fig3 <- ggplot(df1, aes(x=extraversion, y=info_seeking))
fig3 + geom_point() + geom_smooth(method = "lm")

##### AGREEABLENESS AND FUNCTIONS: CORRELATIONS #####
cor.test(df1$agreeableness, df1$interacting)$estimate  #sig
cor.test(df1$agreeableness, df1$commercial)$estimate  #N.S.
cor.test(df1$agreeableness, df1$relaxation)$estimate  #N.S
cor.test(df1$agreeableness, df1$entertainment)$estimate  #N.S
cor.test(df1$agreeableness, df1$info_seeking)$estimate  #N.S

##### AGREEABLENESS AND FUNCTIONS: VISUALIZING THE SIGNIFICANT RELATIONSHIPS #####
fig4 <- ggplot(df1, aes(x=agreeableness, y=interacting))
fig4 + geom_point() + geom_smooth(method = "lm")

##### OPENNESS AND FUNCTIONS #####
cor.test(df1$openness, df1$interacting)$estimate  #sig
cor.test(df1$openness, df1$commercial)$estimate  #sig
cor.test(df1$openness, df1$relaxation)$estimate  #N.S.
cor.test(df1$openness, df1$entertainment)$estimate  #N.S
cor.test(df1$openness, df1$info_seeking)$estimate  #sig

##### OPENNESS AND FUNCTIONS: VISUALIZING THE SIGNIFICANT RELATIONSHIPS #####

fig5 <- ggplot(df1, aes(x = openness, y = interacting))
fig5 + geom_point() + geom_smooth(method = 'lm')

fig6 <- ggplot(df1, aes(x = openness, y = commercial))
fig6 + geom_point() + geom_smooth(method = 'lm')

fig7 <- ggplot(df1, aes(x = openness, y = info_seeking))
fig7 + geom_point() + geom_smooth(method = 'lm')

##### NEUROTICISM AND FUNCTIONS:CORRELATIONS #####
cor.test(df1$neuroticism, df1$interacting)$estimate  #N.S.
cor.test(df1$neuroticism, df1$commercial)$estimate  #N.S.
cor.test(df1$neuroticism, df1$relaxation)$estimate  #sig
cor.test(df1$neuroticism, df1$entertainment)$estimate  #sig
cor.test(df1$neuroticism, df1$info_seeking)$estimate  #N.S

##### NEUROTICISM AND FUNCTIONS: VISUALIZING THE SIGNIFICANT RELATIONSHIPS #####
fig8 <- ggplot(df1, aes(x = neuroticism, y = relaxation))
fig8 + geom_point() + geom_smooth(method = 'lm')

fig9 <- ggplot(df1, aes(x = neuroticism, y = entertainment))
fig9 + geom_point() + geom_smooth(method = 'lm')

##### CONSCIENTIOUSNESS AND FUNCTIONS: CORRELATIONS #####
cor.test(df1$conscientiousness, df1$interacting)$estimate  #N.S.
cor.test(df1$conscientiousness, df1$commercial)$estimate  #N.S.
cor.test(df1$conscientiousness, df1$relaxation)$estimate  #negative sig
cor.test(df1$conscientiousness, df1$entertainment)$estimate  #N.S.
cor.test(df1$conscientiousness, df1$info_seeking)$estimate  #N.S.

##### CONSCIENTIOUSNESS AND FUNCTIONS: VISUALIZING THE SIGNIFICANT RELATIONSHIPS #####
fig10 <- ggplot(df1, aes(x = conscientiousness, y = relaxation))
fig10 + geom_point() + geom_smooth(method = 'lm')

```

### Second Hypothesis: News consumption x Polarization
```{r}
cor.test(df1$news_socialmedia, df1$polarization)
# the correlation between using social media for news consumption and polarization
# was not significant.

# However, the correlation was significant for other news sources (newspaper and cable tv):
cor.test(df1$news_newspaper, df1$polarization)$estimate
cor.test(df1$news_cabletv, df1$polarization)$estimate

# So maybe there is an overall relationship between news consumption and polarization?
# Let's see:
cor.test(df1$news_cons_freq, df1$polarization)
# Yes, that seems to be the case.

fig11 <- ggplot(df1, aes(x = news_cons_freq, y = polarization))
fig11 + geom_point() + geom_smooth(method = 'lm')
```

### Third Hypothesis: Dispositional Traits x Overall Social Media Use
```{r}
# 1) correlations with single variables

cor.test(df1$total_use, df1$boredom) #sig
cor.test(df1$total_use, df1$loneliness) #N.S.
cor.test(df1$total_use, df1$self_esteem) #N.S

lm1 <- lm(total_use ~ boredom, data = df1)
summary(lm1)

lm2 <- lm(total_use ~ boredom + loneliness, data = df1)
summary(lm2)

lm3 <- lm(total_use ~ boredom + loneliness + self_esteem, data = df1)
summary(lm3)

# Are there interaction effects?
summary(lm(total_use ~ boredom*loneliness, data = df1)) #N.S
summary(lm(total_use ~ boredom*self_esteem, data = df1)) #N.S
summary(lm(total_use ~ loneliness*self_esteem, data = df1)) #N.S.

# Only boredom helped predict social media use.

fig12 <- ggplot(df1, aes(x = total_use, y = boredom))
fig12 + geom_point() + geom_smooth(method = 'lm')
```

### Fourth Hypothesis: Well being and active x passive use
Passive uses:
```{r}
# Let's go step by step

# Sanity check: is satisfaction with life predicted by boredom, loneliness, and self-esteem?
lm4 <- lm(sats_w_life ~ boredom + loneliness + self_esteem, data = df1)
summary(lm4)
# Yes, and that's expected. 

# 1) Is satisfaction with life predicted by social media use (overall time)?
lm5 <- lm(sats_w_life ~ total_use, data = df1)
summary(lm5)
# No, and not expected (according to previous findings)

# 2) Is satisfaction with life predicted by smartphone use?
summary(lm(sats_w_life ~ smartphone_SM_daily, data = df1))
# No.

# 3) Is satisfaction with life predicted by passive social media use?
# Passive uses: relaxation.
# Uncertain: entertainment, info_seeking
# Active uses: interacting or commercial. 

# a) relaxation
summary(lm(sats_w_life ~relaxation, data = df1))
# Using social media to relax (distraction) predicts negative changes in well-being.

# b) relaxation and entertainment
summary(lm(sats_w_life~relaxation + entertainment, data = df1))
# Using social media for entertainment as a variable didn't help in our model.

# c) relaxation and info_seeking
summary(lm(sats_w_life~relaxation + info_seeking, data = df1))
# Yeah, it seems that sats with life is predicted negatively by relaxation and positively
# by information seeking.

# d) relaxation and commercial ends
summary(lm(sats_w_life~relaxation + commercial, data = df1))
# Yes. Negative for sats with life and positive for commercial ends.

# e) relaxation, info_seeking, commercial ends
summary(lm(sats_w_life~relaxation + info_seeking + commercial, data = df1))
# relaxation and info_seeking remained significant, but not commercial.
```

 Is well-being predicted by active social media use?
``` {r}
# a) interacting
summary(lm(sats_w_life ~interacting, data = df1))
# yes, positively by interacting on social media.

# b) commercial
summary(lm(sats_w_life ~ commercial, data = df1))
# yes, positively by using it for commercial ends (self-promotion).

# c) interacting and commercial
summary(lm(sats_w_life ~interacting + commercial, data = df1))
# nope, both interacting lost significance when commercial was added.

##### Testing full models to predict well-being based on functions of uses #####

cor.test(df1$interacting, df1$relaxation)
# interacting and relaxation are moderately correlated positively

summary(lm(sats_w_life~relaxation + interacting, data = df1))
# in this model, relaxation (using social media as coping mechanism) predicted
# sats with life negatively while interacted predicted it positively! which is
# surprising, given that they are positively correlated

summary(lm(sats_w_life~relaxation + interacting + info_seeking, data = df1))
# adding info_seeking didn't add much to the variance explained, but was also significant.

# Is there an interaction effect between the two main predictors?
summary(lm(sats_w_life~relaxation*interacting, data=df1))
# No. 

# Let's plot it!
# Preparing the data to plot the multiple regression: 

a <- df1 %>% 
  select(sats_w_life,relaxation) %>% 
  mutate(variable = 'relaxation') %>% 
  dplyr::rename(
    score = relaxation
  )

b <- df1 %>% 
  select(sats_w_life, interacting) %>% 
  mutate(variable = 'interacting') %>% 
  dplyr::rename(
    score=interacting
  )
c <- rbind(a, b)
summary(c)
c$variable <- as.factor(c$variable)

fig13 <- ggplot(c, aes(y = sats_w_life, x = score, color = variable))
fig13 + geom_point() + geom_smooth(method = 'lm')

```


### Testing for possible mediators or moderators
```{r}
# Satisfaction and and interacting (moderated by sense of belonging)
summary(lm(sats_w_life ~ interacting + sense_community_online, data = df1))
# yeah, the effect seems to be mediated by sense of community. 

summary(lm(sats_w_life~relaxation + neuroticism, data = df1))
# yeah, the effect seems to be mediated by neuroticism

summary(lm(sats_w_life ~ interacting + agreeableness, data = df1))
# yeah, the effect seems to be mediated by agreeableness

# what about interacting between functions and usage time?
summary(lm(sats_w_life~ interacting*total_use, data = df1))
# Not significant

# let's redo the models controling for age and gender:
summary(lm(sats_w_life ~ relaxation + interacting + age, data = df1))
# no effect for age

df1$Sex <- recode(df1$Sex, Male = 1, Female = 2)
summary(lm(sats_w_life ~ relaxation + interacting + Sex, data = df1))
# no effect for gender

```






